---
title: "Data Analysis Assignment #2 (75 points total)"
author: "Patel, Dhaval"
output:
  html_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```

### Instructions


Submit both the .Rmd and .html files for grading. You may remove the instructions and example problem above, but do not remove the YAML metadata block or the first, "setup" code chunk.  Address the steps that appear below and answer all the questions. Be sure to address each question with code and comments as needed.  You may use either base R functions or ggplot2 for the visualizations.

-----

##Data Analysis #2

```{r analysis_setup1, message = FALSE, warning = FALSE}

# Perform the following steps to start the assignment.
 
# 1) Load/attach the following packages via library():  flux, ggplot2, gridExtra, moments, rockchalk, car.
# NOTE:  packages must be installed via install.packages() before they can be loaded.

library(dplyr)
library(flux)
library(ggplot2)
library(gridExtra)
library(knitr)
library(rockchalk)
library(tidyverse)

# 2) Use the "mydata.csv" file from Assignment #1 or use the file posted on the course site.  Reading
# the files into R will require sep = "" or sep = " " to format data properly.  Use str() to check file
# structure.

mydata <- read.csv("mydata.csv", sep = ",", stringsAsFactors = TRUE)
# mydata <- read.csv(file.path("c:...", "mydata.csv"), sep = ",")
# mydata <- read.csv(file.path("c:/Rabalone/", "mydata.csv"), sep = ",")

str(mydata)

```

### Test Items starts from here - There are 10 sections - total of 75 points ##############

***#### Section 1: (5 points) ####***

(1)(a) Form a histogram and QQ plot using RATIO. Calculate skewness and kurtosis using 'rockchalk.' Be aware that with 'rockchalk', the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_1a, fig.width = 12}

par(mfrow=c(1,2))
hist(mydata$RATIO, main="Histogram of RATIO", xlab="RATIO", col='red')
qqnorm(mydata$RATIO, main='Q-Q Plot of RATIO', col='purple')
qqline(mydata$RATIO)

rockchalk.skewness <- rockchalk::skewness(mydata$RATIO)
rockchalk.kurtosis <- rockchalk::kurtosis(mydata$RATIO)


cat("Skewness of RATIO:", rockchalk.skewness, "\n")
cat("Kurtosis of RATIO:", rockchalk.kurtosis)


```

(1)(b) Tranform RATIO using *log10()* to create L_RATIO (Kabacoff Section 8.5.2, p. 199-200). Form a histogram and QQ plot using L_RATIO. Calculate the skewness and kurtosis. Create a boxplot of L_RATIO differentiated by CLASS.

```{r Part_1b, fig.width = 12, fig.height = 8}


par(mfrow= c(1,2))
mydata$L_RATIO <- log10(mydata$RATIO)
hist(mydata$L_RATIO, main = "Histogram of L_RATIO", xlab = "L_RATIO", col="cornflowerblue")

qqnorm(mydata$L_RATIO, main = "Q-Q Plot of L_RATIO", col='red')
qqline(mydata$L_RATIO, col='cornflowerblue')

boxplot(mydata$L_RATIO ~ mydata$CLASS, main = "L_Ratio Boxplot",
ylab = "L_Ratio", xlab = "CLASS", beside = TRUE, col = 'cornflowerblue')

rockchalk.skewness2 <- rockchalk::skewness(mydata$L_RATIO)
rockchalk.kurtosis2 <- rockchalk::kurtosis(mydata$L_RATIO)


cat("Skewness of L_RATIO:", rockchalk.skewness2, "\n")
cat("Kurtosis of L_RATIO:", rockchalk.kurtosis2)



```

(1)(c) Test the homogeneity of variance across classes using *bartlett.test()* (Kabacoff Section 9.2.2, p. 222). 

```{r Part_1c}
bartlett.test(L_RATIO ~ CLASS, data = mydata)

```

**Essay Question: Based on steps 1.a, 1.b and 1.c, which variable RATIO or L_RATIO exhibits better conformance to a normal distribution with homogeneous variances across age classes?  Why?** 


***The L_RATIO histogram indicates more normally distributed than RATIO variable is more right positively skewed histogram plot. ***

***Also, L_RATIO has skewness slightly negative of -0.09 but closer than to 0 compare to RATIO skewness value showing L_RATIO is closer to normal distribution. Similarly, L_RATIO has kurtosis value of 0.53 which less than RATIO kurtosis value of 1.66 showing Ratio in the QQ-Plot deviates from normality which is confirmed by the skewness and kurtosis ***

***L_Ratio fails to reject the null hypothesis with the Bartlett test which determines that L_RATIO variance is homogeneous across each age classes***


-----


***#### Section 2 (10 points) ####***

(2)(a) Perform an analysis of variance with *aov()* on L_RATIO using CLASS and SEX as the independent variables (Kabacoff chapter 9, p. 212-229). Assume equal variances. Perform two analyses. First, fit a model with the interaction term CLASS:SEX. Then, fit a model without CLASS:SEX. Use *summary()* to obtain the analysis of variance tables (Kabacoff chapter 9, p. 227).

```{r Part_2a}
summary(aov(L_RATIO ~ CLASS + SEX + CLASS*SEX, data=mydata))
summary(aov(L_RATIO ~ CLASS + SEX, data=mydata))

```

**Essay Question:  Compare the two analyses.  What does the non-significant interaction term suggest about the relationship between L_RATIO and the factors CLASS and SEX?**

***Answer: The interaction between CLASS and SEX is insignificant when compared together there is no strong predictive value between L_RATIO based on the p-value of 0.86 in one way analysis variance test***

***When CLASS and SEX are compared individually have strong predicated values are statistically significant with L_Ratio ***


(2)(b) For the model without CLASS:SEX (i.e. an interaction term), obtain multiple comparisons with the *TukeyHSD()* function. Interpret the results at the 95% confidence level (*TukeyHSD()* will adjust for unequal sample sizes). 

```{r Part_2b}
TukeyHSD(aov(L_RATIO ~ CLASS + SEX, data = mydata), conf.level = 0.95)
```

**Additional Essay Question:  first, interpret the trend in coefficients across age classes. What is this indicating about L_RATIO?  Second, do these results suggest male and female abalones can be combined into a single category labeled as 'adults?' If not, why not?**

***Answer: When analyzing the p-values in Tukey HSD Test at 95% confidence level for the CLASS pairwise comparison shows statistical significances between all age classes except class A2-A1 indicates no significant difference which suggests that these two age classes A1 and A2 could be combined in one age class for better significant difference and relationship with L_RATIO. ***

***Similarly, looking at the second pairwise comparison based on SEX variable, the p-value for Infant and Male (M-I) or Infant and Female (I-F) rejects the null hypothesis.***

***But the comparison between Male and Female (M-F) have no significant difference. This indicates that male and female abalones can be combined into a single category label as adults.***

------

***####  Section 3: (10 points) ####***

(3)(a1) Here, we will combine "M" and "F" into a new level, "ADULT". The code for doing this is given to you. For (3)(a1), all you need to do is execute the code as given.

```{r Part_3a1}
# Here, we show how to define the new variable TYPE using only base R functions:

mydata$TYPE <- factor(ifelse(mydata$SEX == "I", "I", "ADULT"))
table(mydata$TYPE)


```

(3)(a2)  Present side-by-side histograms of VOLUME. One should display infant volumes and, the other, adult volumes. 

```{r Part_3a2, fig.width = 12}
par(mfrow = c(1,2))

hist(mydata$VOLUME[mydata$TYPE == "I"], xlab = "VOLUME", main = "Infant Volume", col = 'coral2')
hist(mydata$VOLUME[mydata$TYPE == "ADULT"], xlab = "VOLUME", main = "Adult Volume", col = 'deepskyblue2')

```


**Essay Question: Compare the histograms.  How do the distributions differ? Are there going to be any difficulties separating infants from adults based on VOLUME?**

***Answer: When comparing histograms, based on the volume Infant distribution is right positively skewed, compare to Adult distribution is more normally distributed. ***

***In addition, the Infant has majority of their volume less than 300 cm^3 whereas the Adult histogram shows the majority of volume between 100 to 600 cm^3. ***

***This indicates there might be some difficulty in categorizing abalones into infant and adults group based on the volume, because we can see there is some overlap in volume around 100 to 300 cm^3 between infants and adults ***

------

(3)(b) Create a scatterplot of SHUCK versus VOLUME and a scatterplot of their base ten logarithms, labeling the variables as L_SHUCK and L_VOLUME. Please be aware the variables, L_SHUCK and L_VOLUME, present the data as orders of magnitude (i.e. VOLUME = 100 = 10^2 becomes L_VOLUME = 2). Use color to differentiate CLASS in the plots. Repeat using color to differentiate by TYPE. 

```{r Part_3b, fig.width = 12, fig.height = 8}
mydata$L_SHUCK <- log10(mydata$SHUCK)
mydata$L_VOLUME <- log10(mydata$VOLUME)

scatter_vol_shuck.class <- ggplot(data = mydata, aes(x = VOLUME, y = SHUCK)) + 
    geom_point(aes(fill = CLASS, colour=CLASS), shape = 16, alpha = 0.50) + 
    theme(legend.position = "bottom") + 
    ggtitle("Shuck vs Volume by Class")

scatter_Lvol_Lshuck.class <- ggplot(data = mydata, aes(x = L_VOLUME, y = L_SHUCK)) + 
    geom_point(aes(fill = CLASS, colour=CLASS), shape = 16, alpha = 0.50) + 
    theme(legend.position = "bottom") +
    ggtitle("L_Shuck vs L_Volume  by Class")

scatter_vol_shuck.type <- ggplot(data = mydata, aes(x = VOLUME, y = SHUCK)) + 
    geom_point(aes(fill = TYPE, colour=TYPE), shape = 16, alpha = 0.50) + 
    theme(legend.position = "bottom") +
    ggtitle("Shuck vs Volume by Type")

scatter_lvol_lshuck.type <- ggplot(data = mydata, aes(x = L_VOLUME, y = L_SHUCK)) + 
    geom_point(aes(fill = TYPE, colour=TYPE), shape = 16, alpha = 0.50) + 
    theme(legend.position = "bottom") +
    ggtitle("L_Shuck vs L_Volume by Type")

grid.arrange(scatter_vol_shuck.class, scatter_Lvol_Lshuck.class, 
             scatter_vol_shuck.type, scatter_lvol_lshuck.type, nrow=2)




```

**Additional Essay Question:  Compare the two scatterplots. What effect(s) does log-transformation appear to have on the variability present in the plot?  What are the implications for linear regression analysis? Where do the various CLASS levels appear in the plots? Where do the levels of TYPE appear in the plots?**

***Answer: When Comparing the two scatter plot, based on the VOLUME and SHUCK there are a lot of overlapping data points which makes it difficult to distinguish between CLASS and TYPE scatter plot. Whereas, log transformation scatter plot reduces variability between L_SHUCK and L_VOLUME, which makes it easier to differentiate between CLASS and TYPE levels. ***

***In addition, we see similar trend in logarithmic plot differentiated by TYPE where Adult abalones tend to be clustered the upper-right quadrants of the graph. The infants have lower shuck and volume in the lower-left to mid quadrant of the plot. It shows more visible correlations between L_SHUCK and L_VOLUME by TYPE indicating linear relationship. ***

***Furthermore, analyzing the log-transformation plot by CLASS it shows that A1 has the lowest SHUCK wight and VOLUME. Also, class A3 and A4 have similar volumes but A3 has higher SHUCK weight. In class A5 older abalones are heavier in weight and have larger volume. All the infants can distinguished in the CLASS A1 in the lower-left quadrant of the plot.***

-----

***####   Section 4: (5 points) ####***

(4)(a1) Since abalone growth slows after class A3, infants in classes A4 and A5 are considered mature and candidates for harvest. You are given code in (4)(a1) to reclassify the infants in classes A4 and A5 as ADULTS. 

```{r Part_4a1}

mydata$TYPE[mydata$CLASS == "A4" | mydata$CLASS == "A5"] <- "ADULT"
table(mydata$TYPE)

```

(4)(a2) Regress L_SHUCK as the dependent variable on L_VOLUME, CLASS and TYPE (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2 and Black Section 14.2). Use the multiple regression model: L_SHUCK ~ L_VOLUME + CLASS + TYPE. Apply *summary()* to the model object to produce results.

```{r Part_4a2}
model <- lm(L_SHUCK ~ L_VOLUME + CLASS + TYPE, data = mydata)
summary(model)


```

**Essay Question:  Interpret the trend in CLASS levelcoefficient estimates? (Hint:  this question is not asking if the estimates are statistically significant. It is asking for an interpretation of the pattern in these coefficients, and how this pattern relates to the earlier displays).**

***Answer: Interpreting linear regression model we can see that the trend as CLASS level increases from A2 to A5 it shows that estimate coefficients decreases in L_SHUCK. This makes sense based on prior observation in scatter plot as L_SHUCK has more significant increase at lower classes.In addition, as class increases from  CLASS A2 to A5 have negative coefficients estimates showing weak negative correlation between L_SHUCK and CLASS has abalones gets older. Also, L_VOLUME has the highest positive weight has a strong effect on L_SHUCK.***

***Hence, based on the negative coefficient estimates it means that the growth in age and SHUCK weight slows down while VOLUME still increases at faster rate.  ***

-----

**Additional Essay Question:  Is TYPE an important predictor in this regression? (Hint:  This question is not asking if TYPE is statistically significant, but rather how it compares to the other independent variables in terms of its contribution to predictions of L_SHUCK for harvesting decisions.)  Explain your conclusion.**

***Answer:TYPE is not an important predictor in this regression because it does not have strong coefficient estimate to predict L_SHUCK compare to CLASS from A3 to A5. Even though TYPE is a fairly significant predictor slightly better compare to CLASS A2 acting as small contributor to predict L_SHUCK based on coefficient estimate of -0.021.***

***The CLASS independent variables from A3 to A5 are more likely to contribute to predictions of L_SHUCK for harvesting decision.***

-----

The next two analysis steps involve an analysis of the residuals resulting from the regression model in (4)(a) (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2).

-----

***#### Section 5: (5 points) ####***

(5)(a) If "model" is the regression object, use model$residuals and construct a histogram and QQ plot. Compute the skewness and kurtosis. Be aware that with 'rockchalk,' the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_5am, fig.width = 12}
par(mfrow = c(1,2))
hist(model$residuals, main="Histogram of Residuals", col="brown", xlab="Residuals")
qqnorm(model$residuals, main="QQ Plot of Residuals", col="brown")
qqline(model$residuals, col="darkgoldenrod1", lwd="2")

rockchalk.skewness_residuals <- rockchalk::skewness(model$residuals)
rockchalk.kurtosis_residuals <- rockchalk::kurtosis(model$residuals)

cat("Skewness of Residuals:", rockchalk.skewness_residuals, "\n")
cat("Kurtosis of Residuals:", rockchalk.kurtosis_residuals)



```

(5)(b) Plot the residuals versus L_VOLUME, coloring the data points by CLASS and, a second time, coloring the data points by TYPE. Keep in mind the y-axis and x-axis may be disproportionate which will amplify the variability in the residuals. Present boxplots of the residuals differentiated by CLASS and TYPE (These four plots can be conveniently presented on one page using *par(mfrow..)* or *grid.arrange()*. Test the homogeneity of variance of the residuals across classes using *bartlett.test()* (Kabacoff Section 9.3.2, p. 222).  

```{r Part_5b, fig.width = 12, fig.height = 8}

mydata$RESIDUALS <- model$residuals

gridExtra::grid.arrange(
  ggplot(mydata, aes(L_VOLUME, RESIDUALS)) + 
    ggtitle("Residuals vs L_VOLUME By Class ") + 
    geom_point(aes(fill = CLASS, colour=CLASS), shape = 16) + theme(legend.position="bottom"),
  
  ggplot(mydata, aes(L_VOLUME, RESIDUALS)) + 
    ggtitle("Residuals vs L_VOLUME By Type ") + 
    geom_point(aes(fill = TYPE, colour=TYPE), shape = 16)+   theme(legend.position="bottom"),
  
  ggplot(mydata, aes(CLASS, RESIDUALS)) +
    ggtitle("Residuals vs Class  Box Plot") +
    geom_boxplot(aes(fill = CLASS)) + theme(legend.position="none"),
  
  ggplot(mydata, aes(TYPE, RESIDUALS)) +
    ggtitle("Residuals vs Type  Box Plot") +
    geom_boxplot(aes(fill = TYPE)) + theme(legend.position="none")
  
)

bartlett.test(RESIDUALS ~ CLASS, data = mydata)

```

**Essay Question:  What is revealed by the displays and calculations in (5)(a) and (5)(b)? Does the model 'fit'?  Does this analysis indicate that L_VOLUME, and ultimately VOLUME, might be useful for harvesting decisions? Discuss.**  

***Answer: The residuals histogram is evenly normally distributed similarly the QQ plot has data points close the trendline as well slightly heavily tails. It can further prove based on skewness and kurtosis are relatively low close to 0 as well***

***The scatterplot and box have data points close to 0, and there are no patterns in the residual values when categorized by CLASS or TYPE. It appears to be healthy we can assume that model is good fit.***

***The VOLUME might be useful harvesting decisions based on the residuals scatterplot categorized by CLASS and TYPE. It indicates large cluster and wide distribution of data points to the right of plot greater than 2 that the sample consists of more heavier volume and older abalones. ***


-----

Harvest Strategy:

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. a specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.The Management needs to make a decision to implement 1 rule that meets the business goal.

The next steps in the assignment will require consideration of the proportions of infants and adults harvested at different cutoffs. For this, similar "for-loops" will be used to compute the harvest proportions. These loops must use the same values for the constants min.v and delta and use the same statement "for(k in 1:10000)."  Otherwise, the resulting infant and adult proportions cannot be directly compared and plotted as requested. Note the example code supplied below.

-----

***#### Section 6: (5 points) ####***

(6)(a) A series of volumes covering the range from minimum to maximum abalone volume will be used in a "for loop" to determine how the harvest proportions change as the "cutoff" changes. Code for doing this is provided.

```{r Part_6a}

idxi <- mydata$TYPE == "I"
idxa <- mydata$TYPE == "ADULT"

max.v <- max(mydata$VOLUME)
min.v <- min(mydata$VOLUME)
delta <- (max.v - min.v)/10000
prop.infants <- numeric(10000)
prop.adults <- numeric(10000)
volume.value <- numeric(10000)

total.infants <- sum(idxi)  
total.adults <- sum(idxa)

for (k in 1:10000) { 
	value <- min.v + k*delta
	volume.value[k] <- value
	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
}

```

(6)(b) Our first "rule" will be protection of all infants. We want to find a volume cutoff that protects all infants, but gives us the largest possible harvest of adults. We can achieve this by using the volume of the largest infant as our cutoff. You are given code below to identify the largest infant VOLUME and to return the proportion of adults harvested by using this cutoff. You will need to modify this latter code to return the proportion of infants harvested using this cutoff. Remember that we will harvest any individual with VOLUME greater than our cutoff.

```{r Part_6b}
# Largest infant volume
(max_inf_vol <- max(mydata$VOLUME[mydata$TYPE == "I"]))  # [1] 526.6383

# Proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > max_inf_vol) /
  total.adults  # [1] 0.2476573

# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > max_inf_vol) /
  total.infants #0

# If we use the largest infant volume, we harvest approximately 24.8% of adults and 0%,
# as expected, of infants.

```

(6)(c) Our next approaches will look at what happens when we use the median infant and adult harvest VOLUMEs. Using the median VOLUMEs as our cutoffs will give us (roughly) 50% harvests. We need to identify the median volumes and calculate the resulting infant and adult harvest proportions for both.

```{r Part_6c}
# Add code to determine the median infant volume:
(med_inf_vol <- median(mydata$VOLUME[mydata$TYPE=="I"]))

# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > med_inf_vol) /
  total.infants

# Add code to calculate the proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_inf_vol) /
  total.adults


# If we use the median infant volume as our cutoff, we harvest almost 50% of our infants
# and a little more than 93% of our adults.


# Add code to determine the median adult volume:
(med_adult_vol <- median(mydata$VOLUME[mydata$TYPE=="ADULT"]))


# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > med_adult_vol) /
  total.infants

# Add code to calculate the proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_adult_vol) /
  total.adults

# If we use the median adult volume as our cutoff, we harvest almost 50% of adults
# and approximately 2.4% of infants.

```

(6)(d) Next, we will create a plot showing the infant conserved proportions (i.e. "not harvested," the prop.infants vector) and the adult conserved proportions (i.e. prop.adults) as functions of volume.value. We will add vertical A-B lines and text annotations for the three (3) "rules" considered, thus far:  "protect all infants," "median infant" and "median adult." Your plot will have two (2) curves - one (1) representing infant and one (1) representing adult proportions as functions of volume.value - and three (3) A-B lines representing the cutoffs determined in (6)(b) and (6)(c).

```{r Part_6d, fig.width = 12, fig.height = 6}
ggplot() +
  geom_line(aes(volume.value, prop.infants, color = "Infants")) + 
  geom_line(aes(volume.value, prop.adults, color = "Adults")) + theme(legend.position = c(0.9, 0.8),legend.title = element_blank()) +   theme(legend.background=element_blank()) + labs(x = "VOLUME", y = "PROPORTION", title = "Proportion of Adults and Infants Protected") + 
  geom_vline(xintercept = med_inf_vol) + annotate("text", label = paste("Median Infant\n", round(med_inf_vol, 1)),  x = med_inf_vol + 58, y = 0.43) + 
  geom_vline(xintercept = med_adult_vol) + annotate("text", label = paste("Median Adult\n", round(med_adult_vol, 1)),  x = med_adult_vol + 58, y = 0.43) + 
  geom_vline(xintercept = max_inf_vol) + annotate("text", label = paste("Largest Infant\n", round(max_inf_vol, 1)),  x = max_inf_vol + 58, y = 0.95) + 
  geom_hline(yintercept = 0.5, linetype=2)
  
  
  
  



```

**Essay Question:  The two 50% "median" values serve a descriptive purpose illustrating the difference between the populations. What do these values suggest regarding possible cutoffs for harvesting?** 

***Answer: The 50% split for infant median values is 133.8 cm^3 is a lot lower than adult median values of 384.6 cm^3 which makes sense as infant have smaller volumes compare to adult abalones. We can see that the adult curve is not as steep as infants which produces a three times as much volume as infants and increasing at faster rate***

***The possible cutoffs for volume of 384 cm^3 criteria should be met for harvesting for atleast 50% of adult abalones.***

-----


More harvest strategies:

This part will address the determination of a cutoff volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. In other words, we want to find the volume value such that the vertical distance between the infant curve and the adult curve is maximum. To calculate this result, the vectors of proportions from item (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

***#### Section 7: (10 points)  ####***

(7)(a) Evaluate a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value. Compare to the 50% "split" points determined in (6)(a). There is considerable variability present in the peak area of this plot. The observed "peak" difference may not be the best representation of the data. One solution is to smooth the data to determine a more representative estimate of the maximum difference.

```{r Part_7a}
diff <- ((1 - prop.adults) - (1 - prop.infants))
head(diff)

```

(7)(b) Since curve smoothing is not studied in this course, code is supplied below. Execute the following code to create a smoothed curve to append to the plot in (a). The procedure is to individually smooth (1-prop.adults) and (1-prop.infants) before determining an estimate of the maximum difference. 

```{r Part_7b}

y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
	family = c("symmetric"))
y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
	family = c("symmetric"))
smooth.difference <- predict(y.loess.a) - predict(y.loess.i)

```

(7)(c) Present a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value with the variable smooth.difference superimposed. Determine the volume.value corresponding to the maximum smoothed difference (Hint:  use *which.max()*). Show the estimated peak location corresponding to the cutoff determined.

Include, side-by-side, the plot from (6)(d) but with a fourth vertical A-B line added. That line should intercept the x-axis at the "max difference" volume determined from the smoothed curve here.

```{r Part_7c, fig.width = 12, fig.height = 6}
max_diff <- volume.value[which.max(smooth.difference)]

diff_plot <- ggplot() + 
  geom_vline(xintercept = max_diff) +
  geom_line(aes(volume.value, diff)) +
  geom_line(aes(volume.value, smooth.difference), linetype=5, color="darkred") +
  labs(x = "Volume", y = "Difference in Proportions Harvested") + 
  annotate("text", label = paste('Volume =', round(max_diff, 3)), x = max_diff + 20, y = 0.35, angle = 90) 

prop <- ggplot() +
  geom_line(aes(volume.value, prop.infants, color = "Infants")) + 
  geom_line(aes(volume.value, prop.adults, color = "Adults")) + theme(legend.position = c(0.9, 0.8),legend.title = element_blank()) + 
  theme(legend.background=element_blank()) + labs(x = "VOLUME", y = "PROPORTION", title = "Proportion of Adults and Infants Protected") + 
  geom_vline(xintercept = med_inf_vol) + annotate("text", label = paste("Med inf \n", round(med_inf_vol, 1)),  x = med_inf_vol + 58, y = 0.43) + 
  geom_vline(xintercept = max_diff) + annotate("text", label = paste("Max diff\n", round(max_diff, 1)),  x = max_diff + 60, y = 0.80) + 
  geom_vline(xintercept = med_adult_vol) + annotate("text", label = paste("Med adl\n", round(med_adult_vol, 1)),  x = med_adult_vol + 70, y = 0.43) + 
  geom_vline(xintercept = max_inf_vol) + annotate("text", label = paste("Largest Infant\n", round(max_inf_vol, 1)),  x = max_inf_vol + 58, y = 0.95) + 
  geom_hline(yintercept = 0.5, linetype="dotted")

grid.arrange(diff_plot, prop, ncol=2)
  

```

(7)(d) What separate harvest proportions for infants and adults would result if this cutoff is used? Show the separate harvest proportions. We will actually calculate these proportions in two ways:  first, by 'indexing' and returning the appropriate element of the (1 - prop.adults) and (1 - prop.infants) vectors, and second, by simply counting the number of adults and infants with VOLUME greater than the vlume threshold of interest.

Code for calculating the adult harvest proportion using both approaches is provided.

```{r Part_7d}

(1 - prop.adults)[which.max(smooth.difference)]  # [1] 0.7416332
# OR,
sum(mydata[mydata$TYPE == "ADULT", "VOLUME"] >
      volume.value[which.max(smooth.difference)]) / total.adults # [1] 0.7416332

(1 - prop.infants)[which.max(smooth.difference)]  # [1] 0.1764706
# OR,
sum(mydata[mydata$TYPE == "I", "VOLUME"] >
      volume.value[which.max(smooth.difference)]) / total.infants # [1] 0.1764706


```

-----

There are alternative ways to determine cutoffs. Two such cutoffs are described below.

-----

***####  Section 8: (10 points)  ####***

(8)(a) Harvesting of infants in CLASS "A1" must be minimized. The smallest volume.value cutoff that produces a zero harvest of infants from CLASS "A1" may be used as a baseline for comparison with larger cutoffs. Any smaller cutoff would result in harvesting infants from CLASS "A1."  

Compute this cutoff, and the proportions of infants and adults with VOLUME exceeding this cutoff. Code for determining this cutoff is provided. Show these proportions. You may use either the 'indexing' or 'count' approach, or both.

```{r Part_8a}

a1.infant <- volume.value[volume.value > max(mydata[mydata$CLASS == "A1" &
        mydata$TYPE == "I", "VOLUME"])][1] # [1] 206.786

cat("Cutoff: ", a1.infant)
cat("\nProportion of Adults with VOLUME exceeding cutoff:", (1 - prop.adults)[which(volume.value == a1.infant)])
cat("\nProportion of Infants with VOLUME exceeding cutoff:",(1 - prop.infants)[which(volume.value == a1.infant)])








```

(8)(b) Next, append one (1) more vertical A-B line to our (6)(d) graph. This time, showing the "zero A1 infants" cutoff from (8)(a). This graph should now have five (5) A-B lines:  "protect all infants," "median infant," "median adult," "max difference" and "zero A1 infants."

```{r Part_8b, fig.width = 12}
ggplot() +
  geom_line(aes(volume.value, prop.infants, color = "Infants")) + 
  geom_line(aes(volume.value, prop.adults, color = "Adults")) + theme(legend.position = c(0.9, 0.8),legend.title = element_blank()) + 
  theme(legend.background=element_blank()) + labs(x = "VOLUME", y = "PROPORTION", title = "Proportion of Adults and Infants Protected") + 
  geom_vline(xintercept = med_inf_vol) + annotate("text", label = paste("Med inf \n", round(med_inf_vol, 1)),  x = med_inf_vol + 30, y = 0.45) + 
  geom_vline(xintercept = a1.infant) + annotate("text", label = paste("Zero \n", "A1 inf \n", round(a1.infant, 1)),  x = a1.infant + 30, y = 0.68) + 
  geom_vline(xintercept = max_diff) + annotate("text", label = paste("Max diff\n", round(max_diff, 1)),  x = max_diff + 60, y = 0.82) + 
  geom_vline(xintercept = med_adult_vol) + annotate("text", label = paste("Med adl\n", round(med_adult_vol, 1)),  x = med_adult_vol + 30, y = 0.45) + 
  geom_vline(xintercept = max_inf_vol) + annotate("text", label = paste("Largest Infant\n", round(max_inf_vol, 1)),  x = max_inf_vol + 58, y = 0.95) + 
  geom_hline(yintercept = 0.5, linetype=2)

```


***#### Section 9: (5 points) ####***

(9)(a) Construct an ROC curve by plotting (1 - prop.adults) versus (1 - prop.infants). Each point which appears corresponds to a particular volume.value. Show the location of the cutoffs determined in (6), (7) and (8) on this plot and label each. 

```{r Part_9, fig.width = 8.5}


cutoffs <- which(volume.value %in% c(max_diff, a1.infant))
infant_cutoffs <- (1 - prop.infants[cutoffs])
adult_cutoffs <- (1 - prop.adults[cutoffs])

cadl_med_adults <- sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_adult_vol) / total.adults
ci_med_adults <- sum(mydata$VOLUME[mydata$TYPE == "I"] > med_adult_vol) / total.infants

cadl_med_inf <- sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_inf_vol) / total.adults
ci_med_inf <- sum(mydata$VOLUME[mydata$TYPE == "I"] > med_inf_vol) / total.infants

cadl_protect_all_inf <- sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > max_inf_vol) / total.adults
ci_protect_all_inf <- sum(mydata$VOLUME[mydata$TYPE == "I"] > max_inf_vol) / total.infants


ggplot(mapping = aes((1 - prop.infants), (1 - prop.adults))) + 
  
  geom_line(color = 'darkblue', size = 1) + 
  geom_abline(intercept = 0, slope = 1, color = 'darkred', linetype = 2) +
  labs(title = "ROC curve of adult and infant harvest proportions", x = "Infant harvest proportion", y = "Adult harvest proportion") + 
  
  geom_point(mapping = aes(infant_cutoffs, adult_cutoffs), shape = 21, size = 5) +
  annotate("text", label = paste('Zero A1 infants \nvol =', round(a1.infant, 1)), x = 0.25, y = 0.90, size = 3) +
  annotate("text", label = paste('Max difference \nvol =', round(max_diff, 1)), x = 0.25, y = 0.68, size = 3) +
  
  geom_point(mapping = aes(round(ci_med_adults,3), round(cadl_med_adults,3)), shape = 21, size = 5) +
  annotate("text", label = paste('Median adult \nvol =', round(med_adult_vol, 1)), x = 0.12, y = 0.46, size = 3) +
  
  geom_point(mapping = aes(ci_med_inf, cadl_med_inf), shape = 21, size = 5) +
  annotate("text", label = paste('Median infant \nvol =', round(med_inf_vol, 1)), x = 0.58, y = 0.90, size = 3) +
  geom_point(mapping = aes(ci_protect_all_inf, cadl_protect_all_inf), shape = 21, size = 5) +
  annotate("text", label = paste('Protect all infants \nvol =', round(max_inf_vol, 1)), x = 0.09, y = 0.20, size = 3)
  

```

(9)(b) Numerically integrate the area under the ROC curve and report your result. This is most easily done with the *auc()* function from the "flux" package.   Areas-under-curve, or AUCs, greater than 0.8 are taken to indicate good discrimination potential. 

```{r Part_9b}
area <- flux::auc(x = (1 - prop.infants), y = (1 - prop.adults))
cat("Area under the curve (AUC): ", area)

```


***#### Section 10: (10 points) ####***

(10)(a) Prepare a table showing each cutoff along with the following:
 	1) true positive rate (1-prop.adults,
 	2) false positive rate (1-prop.infants),
 	3) harvest proportion of the total population

To calculate the total harvest proportions, you can use the 'count' approach, but ignoring TYPE; simply count the number of individuals (i.e. rows) with VOLUME greater than a given threshold and divide by the total number of individuals in our dataset.
 	
```{r Part_10} 	
#1. true positive rate 
true_positive_rate.protect_all_inf <- round(sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > max_inf_vol) / total.adults,3)
true_positive_rate.med_inf <- round(sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_inf_vol) / total.adults,3)
true_positive_rate.med_adults <- round(sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_adult_vol) / total.adults,3)
true_positive_rate.max_diff <- round(sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > max_diff) / total.adults,3)
true_positive_rate.a1_infant <- round(sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > a1.infant) / total.adults,3)

#2. false positive rate
false_positive_rate.protect_all_inf <- round(sum(mydata$VOLUME[mydata$TYPE == "I"] > max_inf_vol) / total.infants,3)
false_positive_rate.med_inf <- round(sum(mydata$VOLUME[mydata$TYPE == "I"] > med_inf_vol) / total.infants,3)
false_positive_rate.med_adults <- round(sum(mydata$VOLUME[mydata$TYPE == "I"] > med_adult_vol) / total.infants,3)
false_positive_rate.max_diff <- round(sum(mydata$VOLUME[mydata$TYPE == "I"] > max_diff) / total.infants,3)
false_positive_rate.a1_infant <- round(sum(mydata$VOLUME[mydata$TYPE == "I"] > a1.infant) / total.infants,3)

#3.Total Population Harvest Proportion
totalharvest_protect_inf <- round(sum(mydata$VOLUME > max_inf_vol) / nrow(mydata),3)
totalharvest_med_inf <- round(sum(mydata$VOLUME > med_inf_vol) / nrow(mydata),3)
totalharvest_med_adult <- round(sum(mydata$VOLUME > med_adult_vol) / nrow(mydata),3)
totalharvest_max_diff <- round(sum(mydata$VOLUME > max_diff) / nrow(mydata),3)
totalharvest_a1_infant <- round(sum(mydata$VOLUME > a1.infant) / nrow(mydata),3)

c1 <- c(round(max_inf_vol,1), true_positive_rate.protect_all_inf, 
        false_positive_rate.protect_all_inf, totalharvest_protect_inf)

c2 <- c(round(med_inf_vol,1), true_positive_rate.med_inf, 
        false_positive_rate.med_inf, totalharvest_med_inf)

c3 <- c(round(med_adult_vol,1), true_positive_rate.med_adults, 
        false_positive_rate.med_adults, totalharvest_med_adult)

c4 <- c(round(max_diff,1), true_positive_rate.max_diff, 
        false_positive_rate.max_diff, totalharvest_max_diff)

c5 <- c(round(a1.infant,1), true_positive_rate.a1_infant, 
        false_positive_rate.a1_infant, totalharvest_a1_infant)

c <- rbind(c1, c2, c3, c4, c5)
colnames(c) <- c("Volume", "TPR", "FPR", "totalHarvest")
rownames(c) <- c("Protect All Infants", "Median Infants", "Median Adults", "Max Difference", "Zero A1 Infants")
kable(c)

```
 	
**Essay Question: Based on the ROC curve, it is evident a wide range of possible "cutoffs" exist. Compare and discuss the five cutoffs determined in this assignment.**   

***Answer: After analyzing the five cutoffs in this assignment, the first cutoff 'Protect All Infants' has lowest proportions of abalones for harvesting with highest volume with 0 false positive rate which makes sense in terms protecting all infants and only harvesting with highest volume***

***The second cutoff 'Median Infants' cutoff has the lowest volume but has the highest proportion of abalones being harvested. This means the following harvesting technique would result into the lowest protection rate of overall abalones. The third cutoff 'Median adults' has the lowest false positive rate, the following cutoff is recommended for harvesting adults abalones since it has decent true positive rate.***

***The fourth cutoff 'Max Difference' has the lowest proportions of both adult and limiting the infant abalones being harvested. The fifth cutoff 'Zero A1 Infants' does make sense the result perspective as we excludes CLASS A1 infants from being harvested the true positive rate increases, but so does the false positive rate.***

------

**Final Essay Question:  Assume you are expected to make a presentation of your analysis to the investigators How would you do so?  Consider the following in your answer:**

1. Would you make a specific recommendation or outline various choices and tradeoffs?
2. What qualifications or limitations would you present regarding your analysis?
3. If it is necessary to proceed based on the current analysis, what suggestions would you have for  implementation of a cutoff?  
4. What suggestions would you have for planning future abalone studies of this type? 

***Answer: 1. The specific recommendation is using the "Protect All Infants' cutoff to help protect infants population since it have 0 false positive rate. Another suggestion is the 'Max Difference' cutoffs has lowest proportion of both infant and abalone group will be helpful to sustain the overall abalones population. I would outline the risks of using the false positive measurements on the cutoffs for harvesting techniques to see effects on the overall population. ***

***2. The limitations that I would present in my analysis is data collection process. The data is manually recorded by researcher based on the type of observation study could lead to human error while counting the rings to determine age. Generally, observation studies not randomized which leads to data not being normally distributed and there were many outliers, such as both residuals scatter plots have large clusters and wide distribution of data to the right of plot greater than 2 this would make it difficult accurately evaluate the abalone classes or type. Furthermore, limits the scope of my analysis***

***3. If it is necessary to proceed based on the current analysis, I would suggest 'Protect all Infants' cutoff preserve infant population that would be especially help during breeding season to sustain overall abalones population growth. Another suggestion would I have is 'Max Difference' as an implementation of cutoff which is lowest proportions of both types, whereas small infant proportion in case would be outliers with high volume this would help to avoid overharvesting***

***4. The suggestion I have for planning on future abalone studies is to take random sample from a larger abalone population. There should be better documentation such as data dictionary with proper description of each variable, harvesting procedure should be improved by implementation of the infants and adults harvested at different cutoffs. To ensure low outliers in terms of abalones sex group and classes less time consuming process should be discovered to get consistent data from sexing abalones and counting shell-rings techniques. Furthermore, researcher could include environmental factors such as pollution, diseases, loss of habitat, and location attributes affecting the proportion of abalones being harvested at different cutoffs in this analysis***
